---
title: "p8105_hw2_sjz2120"
output: github_document
date: "2022-09-28"
---

# Setup
```{r}
library(tidyverse)
library(readxl)
```


# Problem 1

## Data Import and Cleaning
Below we import and clean data from `NYC_Transit_Subway_Entrance_And_Exit_Data.csv`. The process begins with data import, updates variable names, and selects the columns that will be used in later parts fo this problem. We update `entry` from `yes` / `no` to a logical variable. As part of data import, we specify that `Route` columns 8-11 should be character for consistency with 1-7.

```{r}
subway_data =
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
           col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, starts_with("route"),
         entry, exit_only, vending, entrance_type, ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE ))

```


**Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?**

The initial, unmodified subway_data dataset contains [] [] rows x [] columns. It contains [] the variables: [].

So far for data cleaning, I have [].

There are a total of 1868 rows x 19 columns in the resulting dataset which has been cleaned.


Answer the following questions using these data:

* How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St 8th Avenue; 125st Broadway; 125st Lenox); the distinct function may be useful here.
**
* How many stations are ADA compliant?
**
* What proportion of station entrances / exits without vending allow entrance?
**



# Problem 2




# Problem 3