---
title: "p8105_hw2_sjz2120"
output: github_document
date: "2022-09-28"
---

# Setup
```{r}
library(tidyverse)
library(readxl)
```


# Problem 1

## Data Import and Cleaning

Below we import and clean data from `NYC_Transit_Subway_Entrance_And_Exit_Data.csv`. First we use import the data using `read_csv()`, update variable names, and select the columns that will be used in later in this homework. We update the variable `entry` from `yes` / `no` to a logical variable. As part of the data import, we specify that `Route` columns 8-11 should be character for consistency with columns 1-7.

```{r}
subway_data =
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
           col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, starts_with("route"),
         entry, exit_only, vending, entrance_type, ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE ))
```

We still need to "tidy" up subway_data a bit. For instance, route number and route should be variables, so we need to convert all of the `route` variables from wide to long format. This will be useful when focusing on specific routes, but may not be necessary when considering questions that focus on station-level variables. 

The following code chunk selects station name and line, and then uses `distinct()` from `dplyr` to select only unique/distinct rows from the data frame. As a result, the number of rows in this dataset will be equal to the number of unique stations. We see that there are 465 unique stations.

```{r}
subway_data %>%
  select(station_name, line) %>%
  distinct
```

In the next code chunk we now filter according to ADA compliance as an initial step. This produces a dataframe where the number of rows is equal to the number of ADA compliant stations. We see that there are 84 ADA compliant stations.

```{r}
subway_data %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

To compute the proportion of station entrances / exits without vending that allow entrance, we first exclude station entrances that do not allow vending using `filter()`. Then, we focus on the `entry` variable using `pull()`. Since `entry` is a logical variable, we're able to get to the proportion we want by taking the mean of `entry` (since R will treat logical as numeric when taking the mean). We see that the proportion of station entrances / exits without vending that allow entrance is 0.3770492 or about 37.70% of station entrances / exits.

```{r}
subway_data %>%
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

For the last part of Question 1, we need to write a code chunk to identify stations that serve the A train, and to assess how many of these are ADA compliant. First, we tidy the data by converting `route` from wide to long format. Next, we can use tools from previous parts of the question (filtering to focus on the A train, and on ADA compliance; selecting and using `distinct` to obtain dataframes with the required stations in rows). We see that 60 stations serve the A train, and 17 of these stations are ADA compliant.

```{r}
subway_data %>%
    pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

subway_data %>%
    pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```


# Problem 2
This problem uses the Mr. Trash Wheel dataset, available as an Excel file on the course website.

Read and clean the Mr. Trash Wheel sheet:

* specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
* use reasonable variable names
* omit rows that do not include dumpster-specific data
* round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)

```{r}
mr_trash_data =
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
             sheet = "Mr. Trash Wheel", range = "A2:N534") %>%
  janitor::clean_names() %>%
  rename(weight = weight_tons, volume = volume_cubic_yards) %>%
  drop_na(dumpster)

mr_trash_data
```


Use a similar process to import, clean, and organize the data for Professor Trash Wheel, and combine this with the Mr. Trash Wheel dataset to produce a single tidy dataset. To keep track of which Trash Wheel is which, you may need to add an additional variable to both datasets before combining.

```{r}

```


Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in the resulting dataset, and give examples of key variables. For available data, what was the total weight of trash collected by Professor Trash Wheel? What was the total number of sports balls collected by Mr. Trash Wheel in 2020?





# Problem 3